Experiences with planning for natural language generation

### 2008-06-19 trunk/Papers/experiences/introduction.tex
more changes to the introduction

@@

%%%%%
%The purpose of this paper is twofold.

In this paper, we present two recent planning domains that arise in
the context of applying planning to natural language generation.  We
will first review the sentence generation domain, which was cast as a
planning problem by Koller and Stone \citeyear{KolSto07}.  In this
domain, the goal is to generate a single sentence that expresses a
given piece of meaning; the plan encodes this sentence, and the
actions correspond to uttering individual words.  We will then
introduce the GIVE domain (``Generating Instructions in Virtual
Environments'').  GIVE is a new shared task that was recently posed as
a challenge for the NLG community
\cite{alexander07:_shared_task_propos}.  It uses planning as one
module of a NLG system which generates natural-language instructions
that will guide the user towards performing a given task in a virtual
environment.

We will then discuss some experiences with using off-the-shelf
planners in these domains.  In particular, we explore the efficiency
of FF \cite{HoffmannNebel01} and SGPLAN
\cite{hsu06:_new_featur_in_sgplan_for} on instances of our planning
domains of varying sizes.  Our findings are mixed.  On the one hand,
it turns out that modern planners handle the search problems that
arise in NLG quite easily.  On the other hand, they spend tremendous
amounts of time on preprocessing; for instance, FF spends 90\% of its
runtime in the sentence generation domain on grounding out the
literals and actions, and ends up slower than an ad-hoc Java
implementation of GraphPlan that avoids grounding.  We argue that for
domains like ours, which are dominated by the number of actions and the
universe size rather than the combinatorics of the search problem, the
overall runtime of a planner could be improved by grounding more
selectively.

>>>>

The focus of this paper is twofold. We begin by presenting two recent
planning domains that arise in the context of natural language generation:
the sentence generation domain and the GIVE domain. In the sentence
generation task, the goal is to generate a single sentence that expresses a
given meaning. \citeauthor{KolSto07}~\shortcite{KolSto07} cast this domain
as a planning problem, where the plan encodes the necessary sentence and
the actions correspond to uttering individual words. In the GIVE domain
(``Generating Instructions in Virtual Environments''), we describe a new
shared task that was recently posed as a challenge for the NLG community
\cite{alexander07:_shared_task_propos}. GIVE uses planning as one module of
an NLG system that generates natural-language instructions to guide a
user performing a given task in a virtual environment.

We will then discuss some of our experiences using off-the-shelf planners
in these domains. In particular, we explore the efficiency of FF
\cite{HoffmannNebel01} and SGPLAN \cite{hsu06:_new_featur_in_sgplan_for} on
varying-sized instances of our planning domains. We also compare these
results against an ad-hoc Java implementation of GraphPlan.  Our findings
are mixed. On the one hand, it turns out that modern planners handle some
of the search problems that arise in NLG quite easily. (Large problem
instances remain a challenge.) On the other hand, these same planners spend
tremendous amounts of time on preprocessing. For instance, FF spends 90\%
of its runtime in the sentence generation domain on grounding out literals
and actions, and ends up slower than our version of GraphPlan that avoids
grounding. For domains like ours, which are dominated by the number of
actions and the universe size rather than the combinatorics of the search
problem, this observation suggests that the overall runtime of a planner
could be improved by grounding more selectively.

@@

%In this paper, we present two recent planning domains that arise in the
%context of applying planning to natural language generation.  We will
%first review the sentence generation domain, which was cast as a planning
%problem by \citeauthor{KolSto07}~\shortcite{KolSto07}.  In this domain,
%the goal is to generate a single sentence that expresses a given piece of
%meaning; the plan encodes this sentence, and the actions correspond to
%uttering individual words.  We will then introduce the GIVE domain
%(``Generating Instructions in Virtual Environments'').  GIVE is a new
%shared task that was recently posed as a challenge for the NLG community
%\cite{alexander07:_shared_task_propos}.  It uses planning as one module of
%a NLG system which generates natural-language instructions that will guide
%the user towards performing a given task in a virtual environment.

%We will then discuss some experiences with using off-the-shelf planners in
%these domains.  In particular, we explore the efficiency of FF
%\cite{HoffmannNebel01} and SGPLAN \cite{hsu06:_new_featur_in_sgplan_for} on
%instances of our planning domains of varying sizes.  Our findings are
%mixed.  On the one hand, it turns out that modern planners handle the
%search problems that arise in NLG quite easily.  On the other hand, they
%spend tremendous amounts of time on preprocessing; for instance, FF spends
%90\% of its runtime in the sentence generation domain on grounding out the
%literals and actions, and ends up slower than an ad-hoc Java implementation
%of GraphPlan that avoids grounding.  We argue that for domains like ours,
%which are dominated by the number of actions and the universe size rather
%than the combinatorics of the search problem, the overall runtime of a
%planner could be improved by grounding more selectively.
